Modern BTree\\ techniques

### 2020-01-21 btree/presentation.tex
Trie

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-leaf-node/.style={btree-node, fill=greenGood}}
\tikzset{btree-branch-node/.style={btree-node, fill=gray!20}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}
\tikzset{smallbtree-node/.style={btree-node, minimum width=0.4cm, minimum height=0.4cm, rounded corners=0.1cm}}
\tikzset{box/.style={draw, dashed, rounded corners, behind path}}
\tikzset{green-box/.style={box, fill=greenGood}}
\tikzset{blue-box/.style={box, fill=blue!20}}
\tikzset{bloom-filter/.style={
    draw,
    isosceles triangle,
    minimum width=1.2cm,
    shape border rotate=-90,
    isosceles triangle apex angle=60,
    rounded corners
}}
\tikzset{bw-delta/.style={draw, fill=redBad, rounded corners}}
\tikzset{bw-page/.style={draw, fill=greenGood, rounded corners}}
\tikzset{dp-tree-log/.style={draw, minimum width=0.5cm, minimum height=0.5cm, rounded corners=0.1cm}}

@@

        \draw[->, btree-line] (node4) -- (node9);
        \draw[->, btree-line] (node4) -- (node10);
    } {}
}

@@

@@

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Riegger, Christian Vincon, Tobias Petrov, Ilia. (2017).
            Write-optimized indexing with partitioned b-trees. 296-300.
            10.1145/3151759.3151814.
        \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Levandoski, Justin Lomet, David Sengupta, Sudipta. (2012). The Bw-Tree:
        A B-tree for New Hardware Platforms. Proceedings - International
        Conference on Data Engineering. 10.1109/ICDE.2013.6544834.
    \linespread{1.5}

@@

        \onslide<4>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
            \basicleafconnection{}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Zhou, Xinjing Shou, Lidan Chen, Ke Hu, Wei Chen, Gang. (2019).  DPTree:
        differential indexing for persistent memory. Proceedings of the VLDB
        Endowment.
    \linespread{1.5}

>>>>

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-leaf-node/.style={btree-node, fill=greenGood}}
\tikzset{btree-branch-node/.style={btree-node, fill=gray!20}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}
\tikzset{smallbtree-node/.style={btree-node, minimum width=0.4cm, minimum height=0.4cm, rounded corners=0.1cm}}
\tikzset{box/.style={draw, dashed, rounded corners, behind path}}
\tikzset{green-box/.style={box, fill=greenGood}}
\tikzset{blue-box/.style={box, fill=blue!20}}
\tikzset{bloom-filter/.style={
    draw,
    isosceles triangle,
    minimum width=1.2cm,
    shape border rotate=-90,
    isosceles triangle apex angle=60,
    rounded corners
}}
\tikzset{bw-delta/.style={draw, fill=redBad, rounded corners}}
\tikzset{bw-page/.style={draw, fill=greenGood, rounded corners}}
\tikzset{dp-tree-log/.style={draw, minimum width=0.5cm, minimum height=0.5cm, rounded corners=0.1cm}}
\tikzset{trie-node/.style={draw, circle}}

@@

        \draw[->, btree-line] (node4) -- (node9);
        \draw[->, btree-line] (node4) -- (node10);
    } {}

    \ifthenelse{ \equal{#1}{show-fade-connections} } {
        \draw[->, btree-line, color=gray] (node1) -- (node2);
        \draw[->, btree-line, color=gray] (node1) -- (node3);
        \draw[->, btree-line, color=gray] (node1) -- (node4);

        \draw[->, btree-line, color=gray] (node2) -- (node5);
        \draw[->, btree-line, color=gray] (node2) -- (node6);

        \draw[->, btree-line, color=gray] (node3) -- (node7);
        \draw[->, btree-line, color=gray] (node3) -- (node8);

        \draw[->, btree-line, color=gray] (node4) -- (node9);
        \draw[->, btree-line, color=gray] (node4) -- (node10);
    } {}
}

@@

\newcommand{\trienode}[5] {
    \node[trie-node, below=0.35cm of node#1, xshift=#4] (node#2) {};
    \coordinate[above=0.3cm of node#2.center] (node#2-anchor);
    \node[below=0.0cm of node#2-anchor] (node#2-value) {\footnotesize{#3}};
    \draw[-, btree-line] (node#1) -- (node#2);

    \ifthenelse{ \equal{#5}{} } {}
    {
        \node[below=0.05cm of node#2, color=gray] (node#2-total-value) {\tiny{#5}};
    }
}

@@

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Riegger Christian, Vincon Tobias, Petrov Ilia. (2017).
            Write-optimized indexing with partitioned b-trees. 296-300.
        \linespread{1.5}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Levandoski Justin, Lomet David, Sengupta Sudipta. (2012). The Bw-Tree:
        A B-tree for New Hardware Platforms. Proceedings - International
        Conference on Data Engineering.
    \linespread{1.5}

@@

        \onslide<4>
        \begin{tikzpicture}[]
            \basicbtree{show-fade-connections}
            \basicleafconnection{}

@@

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Zhou Xinjing, Shou Lidan, Chen Ke, Hu Wei, Chen Gang. (2019). DPTree:
        differential indexing for persistent memory. Proceedings of the VLDB
        Endowment.
    \linespread{1.5}

    \begin{center}
    \textbf{Trie}
    \vspace{0.2cm}

    \begin{overprint}[12cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \node[minimum width=12cm] (base) {};
            \node[trie-node] (node1) at (base.north) {};
            \trienode{1}{2}{0}{-1.75cm}{}
            \trienode{1}{3}{1}{1.75cm}{}

            \trienode{2}{4}{01}{-1.0cm}{}
            \trienode{2}{5}{1}{1.0cm}{}
            \trienode{3}{6}{0}{-1.0cm}{10}
            \trienode{3}{7}{11}{1.0cm}{}

            \trienode{4}{8}{0}{-0.5cm}{0010}
            \trienode{4}{9}{1}{0.5cm}{}
            \trienode{5}{10}{0}{-0.5cm}{010}
            \trienode{5}{11}{1}{0.5cm}{011}
            \trienode{7}{12}{0}{-0.5cm}{}
            \trienode{7}{13}{1}{0.5cm}{1111}

            \trienode{9}{14}{00}{-0.5cm}{}
            \trienode{9}{15}{00}{0.5cm}{001100}
            \trienode{12}{16}{01}{-0.5cm}{111001}
            \trienode{12}{17}{1}{0.5cm}{11101}

            \trienode{14}{18}{0}{-0.5cm}{0011000}
            \trienode{14}{19}{1}{0.5cm}{0011001}
        \end{tikzpicture}

        \onslide<2>
        \begin{tikzpicture}[]
            \node[minimum width=12cm] (base) {};
            \node[trie-node] (node1) at (base.north) {};
            \trienode{1}{2}{0}{-1.75cm}{}
            \trienode{1}{3}{1}{1.75cm}{}
            \node[draw, dashed, rounded corners, behind path, inner sep=0.2, fit=(node1)(node2)(node3)] {};

            \trienode{2}{4}{01}{-1.0cm}{}
            \trienode{2}{5}{1}{1.0cm}{}
            \trienode{3}{6}{0}{-1.0cm}{10}
            \trienode{3}{7}{11}{1.0cm}{}

            \trienode{4}{8}{0}{-0.5cm}{0010}
            \trienode{4}{9}{1}{0.5cm}{}
            \trienode{5}{10}{0}{-0.5cm}{010}
            \trienode{5}{11}{1}{0.5cm}{011}
            \trienode{7}{12}{0}{-0.5cm}{}
            \trienode{7}{13}{1}{0.5cm}{1111}
            \node[draw, dashed, rounded corners, behind path, inner sep=0.2, fit=(node4)(node5)(node8)(node9)(node10)(node11)] {};
            \node[draw, dashed, rounded corners, behind path, inner sep=0.2, fit=(node6)(node7)(node12)(node13)] {};

            \trienode{9}{14}{00}{-0.5cm}{}
            \trienode{9}{15}{00}{0.5cm}{001100}
            \trienode{12}{16}{01}{-0.5cm}{111001}
            \trienode{12}{17}{1}{0.5cm}{11101}
            \node[draw, dashed, rounded corners, inner sep=0.1, behind path, fit=(node16)(node17)] {};

            \trienode{14}{18}{0}{-0.5cm}{0011000}
            \trienode{14}{19}{1}{0.5cm}{0011001}

            \node[draw, dashed, rounded corners, behind path, inner sep=0.2, fit=(node14)(node15)(node18)(node19)] {};
        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Leis V., Kemper Alfons, Neumann Thomas. (2013). The adaptive radix
            tree: ARTful indexing for main-memory databases. Proceedings -
            International Conference on Data Engineering. 38-49.
        \linespread{1.5}

        \onslide<3>
        \begin{tikzpicture}[]
            \node[minimum width=12cm] (base) {};
            \node[trie-node, fill=greenGood] (node1) at (base.north) {};
            \trienode{1}{2}{0}{-1.75cm, fill=greenGood}{}
            \trienode{1}{3}{1}{1.75cm, fill=greenGood}{}

            \coordinate[left=0.35cm of node2] (node2-shift);
            \node[trie-node, fill=blue!20, below=0.35cm of node2-shift] (node2-shift-target) {};
            \trienode{2-shift-target}{4}{01}{-1.0cm, fill=blue!20}{}
            \trienode{2-shift-target}{5}{1}{1.0cm, fill=blue!20}{}
            \draw[-, btree-line, dotted] (node2) -- (node2-shift-target);

            \coordinate[right=0.35cm of node3] (node3-shift);
            \node[trie-node, fill=yellow!20, below=0.35cm of node3-shift] (node3-shift-target) {};
            \trienode{3-shift-target}{6}{0}{-1.0cm, fill=yellow!20}{10}
            \trienode{3-shift-target}{7}{11}{1.0cm, fill=yellow!20}{}
            \draw[-, btree-line, dotted] (node3) -- (node3-shift-target);

            \trienode{4}{8}{0}{-0.5cm, fill=blue!20}{0010}
            \trienode{4}{9}{1}{0.5cm, fill=blue!20}{}

            \coordinate[right=0.35cm of node5] (node5-shift);
            \node[trie-node, fill=red!20, below=0.35cm of node5-shift] (node5-shift-target) {};
            \trienode{5-shift-target}{10}{0}{-0.5cm, fill=red!20}{010}
            \trienode{5-shift-target}{11}{1}{0.5cm, fill=red!20}{011}
            \draw[-, btree-line, dotted] (node5) -- (node5-shift-target);

            \coordinate[left=0.35cm of node7] (node7-shift);
            \node[trie-node, fill=gray!20, below=0.35cm of node7-shift] (node7-shift-target) {};
            \trienode{7-shift-target}{12}{0}{-0.5cm, fill=gray!20}{}
            \trienode{7-shift-target}{13}{1}{0.5cm, fill=gray!20}{1111}
            \draw[-, btree-line, dotted] (node7) -- (node7-shift-target);

            \coordinate[right=0.35cm of node9] (node9-shift);
            \node[trie-node, fill=Sepia!20, below=0.35cm of node9-shift] (node9-shift-target) {};
            \trienode{9-shift-target}{14}{00}{-0.5cm, fill=Sepia!20}{}
            \trienode{9-shift-target}{15}{00}{0.5cm, fill=Sepia!20}{001100}
            \trienode{14}{18}{0}{-0.5cm, fill=Sepia!20}{0011000}
            \trienode{14}{19}{1}{0.5cm, fill=Sepia!20}{0011001}
            \draw[-, btree-line, dotted] (node9) -- (node9-shift-target);

            \trienode{12}{16}{01}{-0.5cm, fill=gray!20}{111001}
            \trienode{12}{17}{1}{0.5cm, fill=gray!20}{11101}
        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Binna Robert, Zangerle Eva, Pichl Martin, Specht Guenther, Leis
            Viktor. (2018). HOT: A Height Optimized Trie Index \\
            for Main-Memory Database Systems. 521-534.
        \linespread{1.5}
